---
title: "p8105_hw2_ec3663"
author: "Eunsuh Cho"
date: "2023-09-28"
output: github_document
---

```{r setup}
library(tidyverse)
```


# Problem 1

## Cleaning the data

Cleaning the data in pols-month.csv

```{r clean_538_pols}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols = 
  read_csv("data/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 
```

Cleaning the data in snp.csv

```{r clean_538_snp}
snp = 
  read_csv("data/snp.csv") |>
  separate(date, into = c("month", "day", "year"), convert = TRUE) |>
  arrange(year, month) |>
  mutate(month = month.name[month]) |>
  select(year, month, close) 
```

Cleaning the data in unemployment.csv

```{r clean_538_unemp}
unemployment = 
  read_csv("data/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```

## Merging the data

```{r merge_538}
data_538 = 
  left_join(pols, snp) |>
  left_join(x = _, y = unemployment)

str(data_538)
```

## Dataset descriptions

The `pols` data has `r nrow(pols)` observations and `r ncol(pols)` variables and tells us about the party affiliation distribution (democrat or republican) for governors and senators for a given year from years `r range(pols$year)[1]` to `r range(pols$year)[2]`. It also tells us whether the sitting president was a democrat or republican. The `snp` data has `r nrow(snp)` observations and `r ncol(snp)` variables, ranging from years `r range(snp$year)[1]` to `r range(snp$year)[2]`. The `unemployment` data has `r nrow(unemployment)` observations and `r ncol(unemployment)` variables ranging from years `r range(unemployment$year)[1]` to `r range(unemployment$year)[2]`. In Januarys in or after 1975 in which a democrat was president, the **average unemployment rate was `r filter(data_538, month == "January", year >= 1975, president == "dem") |> pull(unemployment) |> mean() |> round(2)`**.  The average unemployment rate over the same time period in which a republican was president was `r filter(data_538, month == "January", year >= 1975, president == "gop") |> pull(unemployment) |> mean() |> round(2)`.



# Problem 2

## Importing and cleaning data

Cleaning Mr. Trash Wheel data

```{r clean_mtw}
library(readxl)

mrtrashwheel_df =
  read_excel("data/trash_data.xlsx", sheet = 1, range = "A2:N586") |> 
  janitor::clean_names() |> 
  mutate(homes_powered = (weight_tons * 500) / 30,
         trash_wheel = 1,
         year = as.numeric(year))
```

Cleaning Professor Trash Wheel data

```{r clean_ptw}
proftrashwheel_df =
  read_excel("data/trash_data.xlsx", sheet = 2, range = "A2:M108") |> 
  janitor::clean_names() |> 
  mutate(homes_powered = (weight_tons * 500) / 30,
         trash_wheel = 2)
```

Cleaning Gwynnda data
```{r clean_gwynnda}
gwynnda_df =
  read_excel("data/trash_data.xlsx", sheet = 4, range = "A2:L157") |> 
  janitor::clean_names() |> 
  mutate(homes_powered = (weight_tons * 500) / 30,
         trash_wheel = 3)
```

## Joining all 3 trash wheel datasets

```{r joining_trash}
trashwheel_df =
  bind_rows(mrtrashwheel_df, proftrashwheel_df, gwynnda_df) |> 
  janitor::clean_names()
```


